---
title: "Leaflet assignment"
author: "Rajaram"
date: "24/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Objective

Create a web page using R Markdown that features a map created with leaflet.

Host your webpage on either GitHub Pages, RPubs or NeoCities.

Your webpage must contain the date that you created the document, and it must contain a map created with leaflet.

## Date

```{r echo=FALSE}
format(Sys.Date(), format= "%d %B, %Y")
```

## Map with Markers

We are going to mark some important tourist destinations in Karnataka in the state of India through markers functionality of leaflet.

```{r markers}
library(leaflet)
kar_tour_places <- data.frame(name = c("Coorg","Mysore","Chikmagalur","Hampi",
                                       "Bangalore", "Udupi", "Murudeshwar",  
                                       "Gokarna", "Bandipur National Park" ,
                                       "Nagarhole National Park",
                                       "Shivanasamudra Falls",
                                       "Sakleshpur","Dandeli",
                                       "Pattadkal","Belur","Jog Falls"),
                              lat = c(12.3375,12.2958,13.3161,
                                      15.3350,12.9716,13.3409,14.0940,
                                      14.5479,11.7788,12.0314,12.2997,
                                      12.9442,15.2361,15.9488,
                                      13.1623,14.2004) ,
                              lng =  c(75.8069,76.6394,75.7720,
                                       76.4600,77.5946,74.7421,74.4899,
                                      74.3188,76.4647,76.1207,77.1773,
                                      75.7866,74.6173,75.8164,
                                      75.8679,74.7922)
                              )

kar_tour_places %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lat=kar_tour_places$lat, lng= kar_tour_places$lng,
             popup = kar_tour_places$name)
```

## Add a cluster
Zoom in the cluster to see individual markers
```{r cluster}
kar_tour_places %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lat=kar_tour_places$lat, lng= kar_tour_places$lng,
             popup = kar_tour_places$name,
             clusterOptions = markerClusterOptions())
```