---
title: "EuStockmarkets using plotly"
author: "Rajaram"
date: "01/11/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

Create a webpage presentation using R markdown that features a plot created with plotly. Host your webpage on either GitHub Pages, R Pubs or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with plotly.

## Date
```{r echo=FALSE}
format(Sys.Date(), format= "%d %B, %Y")
```

## Plot of EuStockmarkets data of diffrent stock indexes.
We are going top use EuStockMarkets data to plot the time series plotting of prices for different indexes.

```{r echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(plotly))
data("EuStockMarkets")
stocks <- as.data.frame(EuStockMarkets) %>%
  gather(index,price) %>%
  mutate(time=rep(time(EuStockMarkets),4))
```

## Plot
```{r warning=FALSE}
plot_ly(stocks,x=~time,y=~price,color=~index,type="scatter",mode="markers") %>%
  layout(title = "Stocks Data")
```





